---
title: ""
author: ""
date: ""
output: 
    bookdown::word_document2: 
        fig_caption: true              
        toc: false                      
        toc_depth: 3                   
        number_sections: true          
        keep_md: false
    bookdown::pdf_document2: 
        toc: true                      
        toc_depth: 3                    
        number_sections: true          
        latex_engine: xelatex
        keep_tex: false
editor_options: 
  chunk_output_type: console
---

```{r setup, message=FALSE,warning=FALSE, echo=FALSE, cache = FALSE, include=TRUE, eval=TRUE}
knitr::opts_chunk$set(echo = FALSE,eval = TRUE,message = FALSE,warning = FALSE,cache=TRUE,include=TRUE,
                      fig.pos = "H", fig.show = "hold",
                      dpi = 200)
# pacotes e dados de entrada
library(magrittr)
library(tidyverse)
library(officer)
library(flextable)
df_ad0 <- read_csv2("G:/CCD/CVE/RESPIRATORIAS/_EQUIPE/DaniloPereiraMori/artigos/artigo_tati_NFcastingSatScan/tab_tudo_RR.csv")

# classificação inicial das predições
f_class <- \(pred, truth) {
  case_when(
    truth == pred & truth == 1 ~ "TP",
    truth == pred & truth == 0 ~ "TN",
    truth != pred & truth == 1 ~ "FP",
    truth != pred & truth == 0 ~ "FN"
  )
}
df_ad <- df_ad0 %>% 
    # classificação binária
    mutate(classRR = ifelse(is.na(RR),0,1)) %>%
    # remoção da predição do limite superior
    filter(id_key %in% c("RR_casos","RR_Mdn","RR_Mdn_mv","RR_obs")) %>%
    # pivot_wider
    select(COD_CIR,data,id,classRR) %>%
    pivot_wider(names_from=id,values_from=classRR) %>%
    # classificação da matriz de confusão:
    mutate(
        across(
            .cols = c(Mdn,Mdn_mv,obs),
            .fns = ~f_class(pred=.x,truth=casos),
            .names = "{.col}_c"
            )
        )

# métricas de classificação de predição
dfsumm <- df_ad %>%
    select(data,ends_with("_c")) %>%
    pivot_longer(ends_with("_c"),
                 names_to = "method",
                 values_to = "classe") %>%
    group_by(data) %>%
    count(method, classe) %>%
    pivot_wider(names_from = classe,
                values_from = n,
                values_fill = 0) %>%
    mutate(
        method = factor(method,
                        levels=c("Mdn_c","Mdn_mv_c","obs_c"),
                        labels=c(
                            "Standard Bayesian prediction",
                            "Moving average of Bayesian predictions",
                            "Naive baseline")),
        Sensitivity = TP/(TP+FN),
        Specificity = TN/(TN+FP),
        PPV = TP/(TP+FP),
        NPV = TN/(TN+FN),
        Accuracy = (TP+TN)/(TP+TN+FP+FN),
        BalancedAccuracy = (Sensitivity + Specificity)/2,
        Youden = Sensitivity + Specificity - 1
    )
write_csv2(dfsumm,file = "G:/CCD/CVE/RESPIRATORIAS/_EQUIPE/DaniloPereiraMori/artigos/artigo_tati_NFcastingSatScan/metricas_classificacao.csv")
```


# Context

SaTScan statistics is the gold standard for territorial classification based on an area's relative risk (RR) compared to the entire region. In our analysis, the area unit corresponds to the CIR and the region to São Paulo State, Brazil.

SaTScan was applied to the consolidated ground-truth data, the naive observations, and two Bayesian predictions (the standard median and the moving average of the median). 

Results were aggregated by epidemiological week and method, yielding four potential prediction outcomes:
- True Positive (TP): The method correctly identifies a significant $\text{RR} > 1$.

- True Negative (TN): The method correctly identifies a significant $\text{RR} \le 1$.

- False Positive (FP): The method incorrectly identifies an $\text{RR} > 1$.

- False Negative (FN): The method incorrectly identifies an $\text{RR} \le 1$.


## Method Performance Evaluation

To evaluate the predictive performance of each method, we calculated standard classification metrics based on the confusion matrix elements (TP, TN, FP, and FN).

The classification analysis primary focuses on four key metrics:

__Sensitivity__ and __Specificity__ measure the method's intrinsic capacity to correctly identify high-risk ($\text{RR} > 1$) and non-high-risk ($\text{RR} \le 1$) areas, respectively.


__Positive Predictive Value (PPV)__ and __Negative Predictive Value (NPV)__ evaluate the post-test reliability of the predictions, quantifying the probability that a generated alert reflects a true elevated risk. 


Additionally, to account for potential class imbalances across epidemiological weeks, we computed overall performance measures, including Accuracy, Balanced Accuracy, and Youden's Index. The mathematical definitions, descriptions, and epidemiological interpretations for all metrics are summarized in Table 1.



```{r tab-descricao-metricas, echo=FALSE}
tb_metrics <- tibble(
  `Metric` = c(
    "Sensitivity",
    "Specificity",
    "Positive Predictive Value (PPV)",
    "Negative Predictive Value (NPV)"#,
    #"Accuracy",
    #"Balanced accuracy",
    #"Youden's index"
  ),

  `Formula` = c(
    "TP / (TP + FN)",
    "TN / (TN + FP)",
    "TP / (TP + FP)",
    "TN / (TN + FN)" #,
    #"(TP + TN) / (TP + TN + FP + FN)",
    #"(Sensitivity + Specificity) / 2",
    #"Sensitivity + Specificity − 1"
  ),

  `Description` = c(
    "Proportion of true positives correctly identified.",
    "Proportion of true negatives correctly identified.",
    "Probability that a positive classification corresponds to a true positive.",
    "Probability that a negative classification corresponds to a true negative."#,
    #"Overall proportion of correctly classified observations.",
    #"Average of sensitivity and specificity.",
    #"Overall ability to discriminate between positive and negative observations."
  ),

  `Interpretation` = c(
    "Ability of the method to detect RR > 1.",
    "Ability of the method to correctly identify RR ≤ 1.",
    "Reliability of positive classifications generated by the method.",
    "Reliability of negative classifications generated by the method."#,
    #"Overall percentage of classifications consistent with the reference.",
    #"Summarizes performance across both classes, even when they are imbalanced.",
    #"Values close to 1 indicate excellent discriminatory performance."
  )
)
#
ft_metricas <- 
  flextable(tb_metrics) |>
  theme_booktabs() |>
  # Tamanho e fonte
  fontsize(size = 10, part = "all") |>
  font(fontname = "Arial", part = "all") |>
  # Alinhamentos
  align(align = "center", part = "header") |>
  align(j = c("Metric", "Description", "Interpretation"),
        align = "left", part = "all") |>
  align(j = "Formula", align = "center", part = "all") |>
  # Cabeçalho em negrito
  bold(part = "header") |>
  # Cor de fundo do cabeçalho (opcional)
  # bg(bg = "#f0f0f0", part = "header") |>
  # Borda inferior do cabeçalho
  # border(border = fp_border(color = "black", width = 1.5), 
  #        part = "header", 
  #        border.bottom = TRUE) |>
  # Largura das colunas
  set_table_properties(layout = "autofit") |>
  # Caption sem bookmark
  set_caption("Table 1: Binary classification performance metrics used to evaluate nowcasting.") %>% 
  autofit()
ft_metricas
```


# Results

## General classification

```{r fig1, include=TRUE,eval=TRUE ,fig.width=10,fig.height=10}
lp <- list()
lp[[1]] <- dfsumm %>% 
    select(data:FN) %>% 
    pivot_longer(cols=FP:FN) %>% 
    mutate(name=factor(name,
                       levels=c("TP","TN","FP","FN"),
                       labels=c(
                           "True Positive (TP)",
                           "True Negative (TN)",
                           "False Positive (FP)",
                           "False Negative (FN)"
                           )
                       )
           ) %>% 
    # mutate(class_name = ifelse(name%in%c("TP","TN"),
    #                            "Correct classifictation",
    #                            "Wrong classifictation"),
    #        class_name = factor(class_name,
    #                            levels=c(
    #                                "Correct classifictation",
    #                                "Wrong classifictation")),
    #        name2 = case_when(
    #            grepl("N",name) ~ "RR < 1",
    #            grepl("P",name) ~ "RR ≥ 1")
    #        ) %>% 
    # ggplot(aes(x=data,y=value,shape=name2,color=name2)) +
    # geom_point() +
    # facet_grid(method~class_name)
    ggplot(aes(x=data,y=value,color=method)) +
    geom_point() +
    geom_line() +
    facet_wrap(~name,nrow=2) +
    labs(y="",x="") +
    theme_bw() +
    theme(legend.position="bottom")
lp[[2]] <- df_ad %>% 
    group_by(data) %>% 
    summarise(SIG = sum(casos==1),nSIG=sum(casos==0)) %>% 
    pivot_longer(-data) %>% 
    mutate(name = factor(name,
                         levels=c("SIG","nSIG"),
                         labels=c(
                             "RR > 1",
                             "RR ≤ 1")
                         ),
           facet_label = "True number of CIR"
           ) %>% 
    ggplot(aes(x=data,y=value,color=name)) +
        geom_point() +
        geom_line() +
        scale_color_manual("",
                           values=c("RR > 1" = "red",
                                    "RR ≤ 1" = "blue")
                           ) +
    labs(y="",x="") +
    facet_wrap(~facet_label) +
    theme_bw() +
    theme(legend.position="bottom")
library(patchwork)
wrap_plots(lp[2:1],ncol=1) +
    plot_annotation(tag_levels = 'A')
```



## Method performance evaluation metrics

```{r resumo-das-metricas,echo=FALSE}
dftab <- dfsumm %>%  
    group_by(method) %>% 
    summarise(
        `Sensitivity availability, n (%)` =
            sprintf("%d (%.1f%%)",
                    sum(!is.na(Sensitivity)),
                    100 * mean(!is.na(Sensitivity))),
        across(
            c(Sensitivity,
              Specificity,
              PPV,
              NPV#,
              #Accuracy,
              #BalancedAccuracy,
              #Youden
              ),
            ~ mean(.x, na.rm = TRUE)
            ),
        .groups = "drop"
        )
ft_metricas <- 
  flextable(dftab) |>
  theme_booktabs() |>
  # Tamanho e fonte
  fontsize(size = 10, part = "all") |>
  font(fontname = "Arial", part = "all") |>
  # Alinhamentos
  align(align = "center", part = "header") |>
  bold(part = "header") |>
  set_table_properties(layout = "autofit") |>
  set_caption("Table 2: Sumamry of the method evaluation performance metrics used to evaluate nowcasting.") %>% 
  autofit()
ft_metricas
```


```{r graficos, include=TRUE,eval=TRUE,fig.width=14,fig.height=10}
dfp <- dfsumm %>% pivot_longer(Sensitivity:Youden) %>% 
    mutate(name=factor(name,
                       levels=c("Sensitivity",
                                "Specificity",
                                "PPV",
                                "NPV",
                                "Accuracy",
                                "BalancedAccuracy",
                                "Youden"))
           ) %>% 
    ungroup()
p <- dfp %>% 
    filter(name %in% c("Sensitivity","Specificity","PPV","NPV")) %>% 
    ggplot(aes(x=data,y=value,color=method)) +
    scale_color_manual("",values=c(
        "Standard Bayesian prediction"="red",
        "Moving average of Bayesian predictions"="blue",
        "Naive baseline"="black")) +
    geom_line(alpha=0.4) +
    geom_point(alpha=0.4) +
    theme_light() +
    labs(y="",x="epidemiological week") +
    facet_wrap(~name,ncol=2) +
    scale_y_continuous(expand=c(0,0.01)) +
    theme(legend.position="bottom",
          strip.text=element_text(color = "black",face = "bold"),
          strip.background = element_rect(fill="white",colour = "black"))
ggsave(filename="G:/CCD/CVE/RESPIRATORIAS/_EQUIPE/DaniloPereiraMori/artigos/artigo_tati_NFcastingSatScan/metricas_clasificacao.jpg",
       plot=p,
       dpi=200,
       width=12,
       height=10)
p
```







```{r}
dfp <- dfsumm %>% pivot_longer(Sensitivity:Youden) %>% 
    mutate(name=factor(name,
                       levels=c("Sensitivity",
                                "Specificity",
                                "PPV",
                                "NPV",
                                "Accuracy",
                                "BalancedAccuracy",
                                "Youden"))
           ) %>% 
    ungroup()
p <- dfp %>% 
    filter(name %in% c("Accuracy","BalancedAccuracy","Youden")) %>% 
    ggplot(aes(x=data,y=value,color=method)) +
    scale_color_manual("",values=c(
        "Standard Bayesian prediction"="red",
        "Moving average of Bayesian predictions"="blue",
        "Naive baseline"="black")) +
    geom_line(alpha=0.4) +
    geom_point(alpha=0.4) +
    theme_light() +
    labs(y="",x="epidemiological week") +
    facet_wrap(~name,ncol=3) +
    scale_y_continuous(expand=c(0,0.01)) +
    theme(legend.position="bottom",
          strip.text=element_text(color = "black",face = "bold"),
          strip.background = element_rect(fill="white",colour = "black"))
```

