---
title: ""
author: ""
date: ""
output: 
    bookdown::word_document2: 
        fig_caption: true              
        toc: false                      
        toc_depth: 3                   
        number_sections: true          
        keep_md: false
    bookdown::pdf_document2: 
        toc: true                      
        toc_depth: 3                    
        number_sections: true          
        latex_engine: xelatex
        keep_tex: false
editor_options: 
  chunk_output_type: console
---

```{r setup, message=FALSE,warning=FALSE, echo=FALSE, cache = FALSE, include=TRUE, eval=TRUE}
knitr::opts_chunk$set(echo = FALSE,eval = TRUE,message = FALSE,warning = FALSE,cache=TRUE,include=TRUE,
                      fig.pos = "H", fig.show = "hold",
                      dpi = 200)
# pacotes e dados de entrada
library(magrittr)
library(tidyverse)
library(readxl)
library(sf)
library(officer)
library(flextable)
####### dados gerais #########
# cir <- cir |> select(Cod6, RS_CIR, COD_CIR)|> rename(CO_MUN_RES=Cod6)
mapa_cir<-read_sf("G:/CCD/CVE/RESPIRATORIAS/05_SRAG_SISTEMA_DE_ALERTA/00_dados_auxiliares/CIR/CIR_UTM.shp")
names(mapa_cir) <- tolower(names(mapa_cir))
mapa_cir <- rename(mapa_cir,cod_space=cod_cir)

# WIS e cobertura
dfwis_cir <- read_csv("dados/ForeScan/Indicadores/indicador_por_municipio.csv") %>% 
  select(-1)
dfwis_sepi <- read_csv("dados/ForeScan/Indicadores/indicador_base.csv") %>% 
  select(-1)

# classificação do satscan
df_ad0 <- read_csv2("dados/tab_tudo_RR.csv")

# classificação inicial das predições
f_class <- \(pred, truth) {
  case_when(
    truth == pred & truth == 1 ~ "TP",
    truth == pred & truth == 0 ~ "TN",
    truth != pred & truth == 1 ~ "FN",
    truth != pred & truth == 0 ~ "FP"
  )
}
df_ad <- df_ad0 %>% 
    # classificação binária
    mutate(classRR = ifelse(is.na(RR),0,1)) %>%
    # remoção da predição do limite superior
    filter(id_key %in% c("RR_casos","RR_Mdn","RR_Mdn_mv","RR_obs")) %>%
    # pivot_wider
    select(COD_CIR,data,id,classRR) %>%
    pivot_wider(names_from=id,values_from=classRR) %>%
    # classificação da matriz de confusão:
    mutate(
        across(
            .cols = c(Mdn,Mdn_mv,obs),
            .fns = ~f_class(pred=.x,truth=casos),
            .names = "{.col}_c"
            )
        )

# métricas de classificação de predição
dfsumm2 <- df_ad %>%
  select(data,ends_with("_c")) %>%
  pivot_longer(ends_with("_c"),
               names_to = "method",
               values_to = "classe") %>%
  group_by(data) %>%
  count(method, classe) %>%
  pivot_wider(names_from = classe,
              values_from = n,
              values_fill = 0) %>%
  mutate(
    method = factor(method,
                    levels=c("Mdn_c","Mdn_mv_c","obs_c"),
                    labels=c(
                        "Standard Bayesian prediction",
                        "Moving average of Bayesian predictions",
                        "Naive baseline")),
    Sensitivity = TP/(TP+FN),
    TruePositiveRate = TP/(TP+FN), # % of real pos correctly identified
    FalsePositiveRate = FP/(FP+TN), # % of real neg incorrectly identified
    Specificity = TN/(TN+FP),
    PPV = TP/(TP+FP),
    PPV = ifelse(is.na(PPV),0,PPV),
    NPV = TN/(TN+FN),
    Accuracy = (TP+TN)/(TP+TN+FP+FN),
    BalancedAccuracy = (Sensitivity + Specificity)/2,
    Youden = Sensitivity + Specificity - 1,
    F1 = 2 * (TruePositiveRate * PPV) / (TruePositiveRate + PPV),
    F1 = ifelse(is.na(F1),0,F1)
    )
#write_csv2(dfsumm,file = "dados/dfsumm2.csv")
df_aud <- inner_join(
  select(dfsumm2,data:Accuracy),
  group_by(df_ad,data) %>% 
    summarise(Real.Pos = sum(casos==1),
              Real.Neg=sum(casos==0)),
  by="data"
)
#
#
#############################################################################
################# padronização de nomes e joins com objetos espaciais #######
#############################################################################
clean_cir <- function(x) {
  x %>%
    # 1. Remove acentos (ex: "Jundiaí" -> "Jundiai")
    stringi::stri_trans_general(id = "Latin-ASCII") %>%
    # 2. Converte tudo para maiúsculas (ex: "Jundiai" -> "JUNDIAI")
    toupper() %>%
    # 3. Remove múltiplos espaços em branco, se houver
    stringr::str_squish()
}
mapa_cir$rs_cir <- clean_cir(mapa_cir$rs_cir)
# incluindo o shapefile nas métricas da predição do modelo bayesiano
## no 
dfwis_cir$cir <- clean_cir(dfwis_cir$cir)
vwis <- unique(dfwis_cir$cir)
vmapa <- unique(mapa_cir$rs_cir)
base::setdiff(vwis,vmapa)


dfwis_cir <- left_join(mapa_cir,dfwis_cir,by=c("rs_cir"="cir"))
#@-------- criação das métricas de confusão por localidade ------@#

dfsumm_cir <- df_ad %>%
  select(COD_CIR,ends_with("_c")) %>%
  pivot_longer(ends_with("_c"),
               names_to = "method",
               values_to = "classe") %>%
  group_by(COD_CIR) %>%
  count(method, classe) %>%
  pivot_wider(names_from = classe,
              values_from = n,
              values_fill = 0) %>%
  mutate(
    method = factor(method,
                    levels=c("Mdn_c","Mdn_mv_c","obs_c"),
                    labels=c(
                        "Standard Bayesian prediction",
                        "Moving average of Bayesian predictions",
                        "Naive baseline")),
    TruePositiveRate = TP/(TP+FN), # % of real pos correctly identified
    FalsePositiveRate = FP/(FP+TN), # % of real neg incorrectly identified
    Specificity = TN/(TN+FP), # taxa de verdadeiro negativo TNR
    PPV = TP/(TP+FP), # % das atribuições positivas que são verdadeiras
    NPV = TN/(TN+FN), # % das atribuiçãos negativas que são verdadeiras
    Accuracy = (TP+TN)/(TP+TN+FP+FN)) %>% # % de certos de todas as atrib
  mutate(
    PPV = ifelse(is.na(PPV),0,PPV), # em casos que há 0 atrib positivas (TP+FP=0)
    TruePositiveRate = ifelse(is.na(TruePositiveRate),0,TruePositiveRate), # casos sem atribuições positivas corretas e sem erros em atribuições negativas, COD_CIR: 35012 35031 35034 35061 35062 35073 35081 35082 35083 35092 35094 35101 35131 35142 35143 35161 35162 35163
    # BalancedAccuracy = (TruePositiveRate + Specificity)/2,
    Youden = TruePositiveRate - FalsePositiveRate,
    F1 = 2 * (TruePositiveRate * PPV) / (TruePositiveRate + PPV)
  )
dfsumm_cir <- mutate(dfsumm_cir,
                     F1 = ifelse(is.na(F1),0,F1))
dfsumm_cir <- left_join(mapa_cir,dfsumm_cir,by=c("cod_space"="COD_CIR"))
```
```{r eval=FALSE,include=FALSE}
df_aud %>% 
  filter(PPV==1,grepl("Naive",method)) %>% 
  select(method:Sensitivity,PPV,Real.Pos,Real.Neg)
dfsumm %>%
    group_by(data,method) %>% 
    mutate(total = sum(FN,TN,TP,FP)) %>% 
    summary()
dfsumm_cir %>% 
    sf::st_drop_geometry() %>% 
    group_by(rs_cir,method) %>% 
    mutate(total = sum(FN,TN,TP,FP)) %>% 
    select(rs_cir:FP,total) %>% 
    filter(grepl("POLO",rs_cir)) %>% 
    mutate(TPR = TP / (TP + FN),
           FPR = FP / (FP + TN)
           )
```

# Context

SaTScan statistics is the gold standard for territorial classification based on an area's relative risk (RR) compared to the entire region. In our analysis, the area unit corresponds to the CIR and the region to São Paulo State, Brazil.

SaTScan was applied to the consolidated ground-truth data, the naive observations, and two Bayesian predictions (the standard median and the moving average of the median). 

Results were aggregated by epidemiological week and method, yielding four potential prediction outcomes:
- True Positive (TP): The method correctly identifies a significant $\text{RR} > 1$.

- True Negative (TN): The method correctly identifies a significant $\text{RR} \le 1$.

- False Positive (FP): The method incorrectly identifies an $\text{RR} > 1$.

- False Negative (FN): The method incorrectly identifies an $\text{RR} \le 1$.


## Method Performance Evaluation

To evaluate the predictive performance of each method, we calculated standard classification metrics based on the confusion matrix elements (TP, TN, FP, and FN).

The classification analysis primary focuses on four key metrics:

__Sensitivity__ and __Specificity__ measure the method's intrinsic capacity to correctly identify high-risk ($\text{RR} > 1$) and non-high-risk ($\text{RR} \le 1$) areas, respectively.


__Positive Predictive Value (PPV)__ and __Negative Predictive Value (NPV)__ evaluate the post-test reliability of the predictions, quantifying the probability that a generated alert reflects a true elevated risk. 


Additionally, to account for potential class imbalances across epidemiological weeks, we computed overall performance measures, including Accuracy, Balanced Accuracy, and Youden's Index. The mathematical definitions, descriptions, and epidemiological interpretations for all metrics are summarized in Table 1.



```{r tab-descricao-metricas, echo=FALSE}
tb_metrics <- tibble(
  `Metric` = c(
    #"Sensitivity",
    "True Positive Rate (TPR)",
    "False Positive Rate (FPR)", # % of real neg incorrectly identified
    "Specificity",
    "Positive Predictive Value (PPV)",
    "Negative Predictive Value (NPV)",
    "F1-score",
    "Accuracy",
    #"Balanced accuracy",
    "Youden's index"
  ),

  `Formula` = c(
    "TP / (TP + FN)", # TPR
    "FP / (FP + TN)", # FPR
    "TN / (TN + FP)", # Spec
    "TP / (TP + FP)", # PPV
    "TN / (TN + FN)" , # NPV
    "2 × (PPV × TPR) / (PPV + TPR)", # F1-score (ou 2 × TP / (2 × TP + FP + FN))
    "(TP + TN) / (TP + TN + FP + FN)", # Accuracy
    #"(Sensitivity + Specificity) / 2", 
    "TPR − FPR" # Youden
  ),

  `Description` = c(
    "Proportion of real positives correctly identified.", # TPR
    "Proportion of real negatives incorrectly identified.", # NPR
    "Proportion of real negatives correctly identified.", #Spec
    "Probability that a positive classification corresponds to a true positive.", #PPV
    "Probability that a negative classification corresponds to a true negative.", # NPV
    "Harmonic mean of precision (PPV) and recall (TPR).", # F1-score
    "Overall proportion of correctly classified observations.", # accur
    #"Average of sensitivity and specificity.",
    "Difference between true alarms and false alarms." # youden
  ),

  `Interpretation` = c(
    "Ability of the method to detect RR > 1.",
    "Tendency of the method to falsely signal RR > 1.",
    "Ability of the method to detect RR ≤ 1.",
    "Reliability of positive classifications generated by the method.",
    "Reliability of negative classifications generated by the method.",
    "Overall performance in detecting high-risk events (RR > 1) while minimizing false alarms.", # F1-score
    "Overall percentage of classifications consistent with the reference.",
    #"Summarizes performance across both classes, even when they are imbalanced.",
    "Values close to 1 indicate excellent discriminatory performance."
  )
)
#
ft_metricas <- 
  flextable(tb_metrics) |>
  theme_booktabs() |>
  # Tamanho e fonte
  fontsize(size = 10, part = "all") |>
  font(fontname = "Arial", part = "all") |>
  # Alinhamentos
  align(align = "center", part = "header") |>
  align(j = c("Metric", "Description", "Interpretation"),
        align = "left", part = "all") |>
  align(j = "Formula", align = "center", part = "all") |>
  # Cabeçalho em negrito
  bold(part = "header") |>
  # Cor de fundo do cabeçalho (opcional)
  # bg(bg = "#f0f0f0", part = "header") |>
  # Borda inferior do cabeçalho
  # border(border = fp_border(color = "black", width = 1.5), 
  #        part = "header", 
  #        border.bottom = TRUE) |>
  # Largura das colunas
  set_table_properties(layout = "autofit") |>
  # Caption sem bookmark
  set_caption("Binary classification performance metrics used to evaluate nowcasting.") %>% 
  autofit()
ft_metricas
```


# Results

## Figure 1 from main text

Model evaluation metrics: cover percentage and WIS decomposition

```{r fig1-main-text,eval=TRUE,include=FALSE}
library(ggspatial)
library(magick)
library(patchwork)
library(cowplot)
f_imgtoggplot <- function(caminho_imagem, largura_px = 2000) {
  img <- image_read(caminho_imagem) %>% 
    image_trim() %>% 
    # Aumenta/padroniza a largura da imagem mantendo a proporção
    image_resize(geometry_size_pixels(width = largura_px))
  
  ggdraw() + draw_image(img)
}
# primeiro cobertura e depois 
lp <- list()
# figura construida pela Tatiana e adaptada aqui:
vsepi <- dfwis_sepi$base %>% unique %>% sort() %>% as.character()
dfp <- dfwis_sepi %>%
  select(base, cob50, cob95) %>% 
  pivot_longer(
    cols = c(cob50, cob95), 
    names_to = "cobertura", 
    values_to = "valor"
  ) %>% 
  mutate(
    sepi = as.character(base),
    sepi = factor(sepi,levels=vsepi),
    cobertura = factor(
      cobertura, 
      levels = c("cob50","cob95"),
      labels = c("50%","95%")
    )
  )
  #f_imgtoggplot("~/Documentos/projeto_DOC/SRAG/artigos/metricas_forescan/pasta_tati_forescan/Indicadores/coverage_base.png")
lp[[1]] <- dfp %>%
  ggplot(aes(x = sepi, y = valor, fill = cobertura)) +
    geom_col(position = "identity",alpha=0.85) +
    scale_fill_manual(
      values = c(
        "50%" = "darkblue",
        "95%" = "#33CCFF" 
      )
    ) +
  geom_hline(yintercept = c(50,95),color="darkred",linetype=2) +
    labs(
      x = "",
      y = "Empirical coverage",
      fill = "Prediction interval"
    ) +
    theme_bw() +
    theme(
      text=element_text(size=10),
      panel.grid.minor = element_blank(),
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1,face="bold"),
      legend.position = "top",
      strip.background = element_rect(fill = "white")
    )
# figura com os mapas por cobertura
lp[[2]] <- dfwis_cir %>% 
  select(starts_with("cob")) %>% 
  pivot_longer(starts_with("cob")) %>%
  mutate(name=factor(name,
                     levels=c("cob50","cob95"),
                     labels=c("Pred. Int. 50%","Pred. Int. 95%"))) %>% 
  ggplot(aes(fill=value)) +
    geom_sf(color="white") +
    scale_fill_viridis_c("Empirical coverage (%)") +
    facet_wrap(~name,ncol=2) +
  # 1. Ajuste dos eixos para não ter sobras (margens zeradas)
    coord_sf(expand = FALSE) +
    # 2. Adiciona a Barra de Escala (canto inferior esquerdo)
    annotation_scale(
      location = "bl", 
      width_hint = 0.3,
      style = "ticks"
    ) +
    # 3. Adiciona o Norte (canto superior direito)
    annotation_north_arrow(
      location = "tr", 
      which_north = "true",
      pad_x = unit(0.15, "in"), 
      pad_y = unit(0.15, "in"),
      height = unit(0.8, "cm"),  # Reduz o tamanho da seta
      width = unit(0.8, "cm"),
      style = north_arrow_orienteering(
        fill = c("black", "white"),
        line_col = "black",
        text_size = 7
        )
      ) +
    # 4. Manutenção do tema limpo para mapas
    theme_bw() +
    theme(
      legend.position="top",
      panel.grid = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(face = "bold",size=12),
      text = element_text(size=10)
    )
lp[[3]] <- f_imgtoggplot("dados/ForeScan/Indicadores/wis_base.png")
lp[[4]] <- dfwis_cir %>% 
  select(dispersao:subprev) %>% 
  pivot_longer(dispersao:subprev) %>% 
  mutate(name = factor(name,
                       levels=c("dispersao","subprev"),
                       labels=c("spread","underprediction"))) %>% 
  ggplot(aes(fill=log(value))) +
    geom_sf(color="white") +
    scale_fill_gradient(
      low = "#568203", 
      high = "red",
      name = "log(WIS component)"
      ) +
    facet_wrap(~name,nrow=1) +
    # 1. Ajuste dos eixos para não ter sobras (margens zeradas)
    coord_sf(expand = FALSE) +
    # 2. Adiciona a Barra de Escala (canto inferior esquerdo)
    annotation_scale(
      location = "bl", 
      width_hint = 0.3,
      style = "ticks"
    ) +
    # 3. Adiciona o Norte (canto superior direito)
    annotation_north_arrow(
      location = "tr", 
      which_north = "true",
      pad_x = unit(0.15, "in"), 
      pad_y = unit(0.15, "in"),
      height = unit(0.8, "cm"),  # Reduz o tamanho da seta
      width = unit(0.8, "cm"),
      style = north_arrow_orienteering(
        fill = c("black", "white"),
        line_col = "black",
        text_size = 7
        )
      ) +
    # 4. Manutenção do tema limpo para mapas
    theme_bw() +
    theme(
      legend.position="top",
      panel.grid = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(face = "bold",size=12),
      text = element_text(size=10)
    )
painel_final <- wrap_plots(
  lp[c(1,3,2,4)], 
  ncol = 2, 
  widths = c(1, 1)
  ) + 
  plot_annotation(tag_levels = 'A')
vpath <- ggsave(filename = "figuras_e_relatorios/fig1_md_perf.png",
                plot=painel_final,
                dpi=200,
                width=14,
                height=10)
library(magick)
f_img <- \(imgpath){
    vimg <- image_read(imgpath) %>%
        image_trim()
    image_write(vimg,imgpath)
}
f_img(vpath)
painel_final
```





## Figure 2 from main text: SaTScan-based risk classification of CIRs

```{r fig2, include=FALSE,eval=TRUE ,fig.width=10,fig.height=14}
lp <- list()
lp[[1]] <- df_ad %>% 
    group_by(data) %>% 
    summarise(SIG = sum(casos==1),nSIG=sum(casos==0)) %>% 
    pivot_longer(-data) %>% 
    mutate(name = factor(name,
                         levels=c("SIG","nSIG"),
                         labels=c(
                             "RR > 1",
                             "RR ≤ 1")
                         ),
           facet_label = "Real number of CIR"
           ) %>% 
    ggplot(aes(x=data,y=value,color=name)) +
        geom_point() +
        geom_line() +
        scale_color_manual("",
                           values=c("RR > 1" = "orange",
                                    "RR ≤ 1" = "darkgreen")
                           ) +
    labs(y="",x="") +
    ggtitle(label="A)") +
    facet_wrap(~facet_label) +
    theme_bw() +
    theme(legend.position="bottom",
          legend.margin = margin(t = -5, b = 0, unit = "pt"),
          legend.box.margin = margin(t = -8, b = 0, unit = "pt"),
          legend.box.spacing = unit(0, "pt"),
          plot.title = element_text(size = 18),
          strip.background = element_rect(fill="white",colour = "black"))
#
nsepi <- length(unique(df_ad$data))
dfp <- df_ad %>% 
    group_by(COD_CIR) %>% 
    summarise(SIG = sum(casos==1)*100/nsepi,nSIG=sum(casos==0)*100/nsepi) %>% 
    pivot_longer(-COD_CIR) %>% 
    inner_join(x=mapa_cir,y=.,by=c("cod_space"="COD_CIR")) %>% 
    mutate(name = factor(name,
                         levels=c("SIG","nSIG"),
                         labels=c(
                             "RR > 1",
                             "RR ≤ 1")
                         )
           )
fplot <- \(vclass){
    vcolor <- case_when(
        grepl(">",vclass)~"orange",
        grepl("≤",vclass)~"darkgreen",
    )
    dfp %>% 
        filter(grepl(vclass,name)) %>% 
        ggplot(aes(fill=value)) +
        geom_sf(color="white") +
        scale_fill_gradient2(low = "darkblue", 
                             high = "red", 
                             mid = "yellow", 
                             midpoint = 50,
                             name = "% of weeks",
                             limits=c(0,100)) +
        facet_wrap(~name,nrow=1) +
        coord_sf(expand = FALSE) +
        # 2. Adiciona a Barra de Escala (canto inferior esquerdo)
        annotation_scale(location = "bl", width_hint = 0.3,style = "ticks") +
        # 3. Adiciona o Norte (canto superior direito)
        annotation_north_arrow(
                    location = "tr", 
                    which_north = "true",
                    pad_x = unit(0.15, "in"), 
                    pad_y = unit(0.15, "in"),
                    height = unit(0.8, "cm"),  # Reduz o tamanho da seta
                    width = unit(0.8, "cm"),
                    style = north_arrow_orienteering(
                        fill = c("black", "white"),
                        line_col = "black",
                        text_size = 7)
                    ) +
        # 4. Manutenção do tema limpo para mapas
        theme_bw() +
        theme(
            legend.position="bottom",
            panel.grid = element_blank(),
            axis.text = element_blank(),
            axis.ticks = element_blank(),
            strip.background = element_rect(fill=vcolor),
            strip.text = element_text(face = "bold",size=8,color="white"),
            text = element_text(size=10)
            )
}
lpi <- lapply(levels(dfp$name),fplot)
lp[[2]] <- wrap_elements(
    wrap_plots(lpi,ncol=2, guides = "collect") +
    plot_annotation(title = "B)") &
    theme(legend.position = "bottom",
          plot.margin = margin(t = 2, r = 2, b = 2, l = 2, unit = "pt"),
          # altura e largura
          legend.key.height = unit(0.2, "cm"), 
          legend.key.width  = unit(0.6, "cm"), 
              # fonte elementos
              legend.text = element_text(size = 5),
              #fonte do título
              legend.title = element_text(size = 6, face = "bold"),
              #aproxima a legenda
              legend.box.margin = margin(t = -5, b = 0, unit = "pt")
    )
)
#
dfp <- dfsumm2 %>% 
  select(data:method,contains("Rate")) %>% 
  pivot_longer(cols=contains("Rate")) %>% 
    mutate(name=factor(name,
                       levels=c("TruePositiveRate",
                                "FalsePositiveRate"
                                ),
                       labels = c("True Positive Rate",
                                  "False Positive Rate"))
           )
fplot <- \(vgrepl,vrange=c(0,1)){
  dfp0 <- dfp %>% filter(grepl(vgrepl,name)) %>% 
    mutate(method=factor(method,
                         levels=c(
                           "Standard Bayesian prediction",
                           "Moving average of Bayesian predictions",
                           "Naive baseline"),
                         labels = c(
                           "Std. Bayesian prediction",
                           "MA Bayesian prediction", 
                           "Naive baseline"
                           )
                         )
                   )
  vtitle <- ifelse(grepl("True",vgrepl),"C)","E)")
  dfp0 %>% 
  ggplot(aes(x=data,y=value,color=method)) +
    scale_color_manual("",values=c(
        "Std. Bayesian prediction"="red",
        "MA Bayesian prediction"="blue",
        "Naive baseline"="black")) +
    geom_line(alpha=0.5,linewidth = 1.5) +
    geom_point(alpha=0.5,size = 1.5) +
    theme_bw() +
    labs(y="",x="") +
    ggtitle(label=vtitle) +
    facet_wrap(~name,ncol=2) +
    scale_y_continuous(expand=c(0,0.01),limits = vrange) +
    theme(legend.position="bottom",
          legend.margin = margin(t = -5, b = 0, unit = "pt"),
          legend.box.margin = margin(t = -8, b = 0, unit = "pt"),
          plot.title = element_text(size = 18),
          strip.text=element_text(color = "black",face = "bold",size=8),
          strip.background = element_rect(fill="white",colour = "black"))
}
lpi_sepi <- lapply(levels(dfp$name),fplot)
lpi_sepi[[3]] <- lp[[1]]
names(lpi_sepi) <- c(levels(dfp$name),"Real")

# por cir
dfp <- dfsumm_cir %>% 
  select(method,contains("Rate")) %>% 
  pivot_longer(contains("Rate")) %>% 
   mutate(name=factor(name,
                       levels=c("TruePositiveRate",
                                "FalsePositiveRate"),
                       labels = c("True Positive Rate",
                                  "False Positive Rate"))
           )
f_ggplot <- \(vgrepl,vrange=c(0,1)){
    dfp0 <- dfp %>% filter(grepl(vgrepl,name))
    vmethods <- levels(dfp$method)
    vfacetlabel <- dfp0$name[1] %>% as.character()
    vtitle <- ifelse(grepl("True",vgrepl),"D)","F)")
    fplot <- \(vmet){
        vcolorstrip <- case_when(
            grepl("Naive",vmet) ~ "black",
            grepl("Standard",vmet) ~ "red",
            grepl("Moving",vmet) ~ "blue"
            )
        dfi <- filter(dfp0,method==vmet) %>% 
            mutate(method=factor(method,
                                 levels=c(
                                     "Standard Bayesian prediction",
                                     "Moving average of Bayesian predictions",
                                     "Naive baseline"),
                                 labels = c(
                                   "Standard Bayes",
                                   "MA Bayes", 
                                   "Naive baseline"
                                   )
                                 )
                   )
        dfi %>% 
            ggplot(aes(fill=value)) +
            geom_sf(color="gray") +
            scale_fill_gradient2(low = "purple", 
                                 mid = "white", 
                                 high = "yellow", 
                                 midpoint = 0.50,
                                 name = vfacetlabel,
                                 limits=vrange) +
            facet_wrap(~method,ncol=1) +
            # 1. Ajuste dos eixos para não ter sobras (margens zeradas)
            coord_sf(expand = FALSE) +
            # 2. Adiciona a Barra de Escala (canto inferior esquerdo)
            annotation_scale(location = "bl", width_hint = 0.3,style = "ticks") +
            # 3. Adiciona o Norte (canto superior direito)
            annotation_north_arrow(
                location = "tr", 
                which_north = "true",
                pad_x = unit(0.15, "in"), 
                pad_y = unit(0.15, "in"),
                height = unit(0.8, "cm"),  # Reduz o tamanho da seta
                width = unit(0.8, "cm"),
                style = north_arrow_orienteering(
                    fill = c("black", "white"),
                    line_col = "black",
                    text_size = 7)
                ) +
            # 4. Manutenção do tema limpo para mapas
            theme_bw() +
            theme(
                legend.position="top",
                panel.grid = element_blank(),
                axis.text = element_blank(),
                axis.ticks = element_blank(),
                strip.background = element_rect(fill = vcolorstrip),
                strip.text = element_text(face = "bold",size=8,color="white"),
                text = element_text(size=10)
            )
    }
    lpi <- lapply(grep("Baye",vmethods,value = TRUE),fplot)
    wrap_elements(
        wrap_plots(lpi,nrow=1, guides = "collect") +
        plot_annotation(title=vtitle) &
        theme(legend.position = "bottom",
              plot.margin = margin(t = 2, r = 2, b = 2, l = 2, unit = "pt"),
              # altura e largura
              legend.key.height = unit(0.2, "cm"), 
              legend.key.width  = unit(0.6, "cm"), 
              # fonte elementos
              legend.text = element_text(size = 5),
              #fonte do título
              legend.title = element_text(size = 6, face = "bold"),
              #aproxima a legenda
              legend.box.margin = margin(t = -5, b = 0, unit = "pt")
              )
    )
}  
lpi_cir <- lapply(levels(dfp$name),f_ggplot)
lpi_cir[[3]] <- lp[[2]]
names(lpi_cir) <- c(levels(dfp$name),"Real")
#
f_ggsave <- \(vpath,pplot,vw,vh){
  ggsave(filename = vpath,
           plot=pplot,
           dpi=200,
           width = vw,
           height = vh)
}
df_save <- data.frame(label=paste0(LETTERS[1:6],")"),
                      vnames=rep(names(lpi_cir)[c(3,1:2)],each=2),
                      vlist=rep(c("_sepi","_cir"),3)) 
lpath <- lapply(split(df_save,df_save$label),\(dfi){
  lp <-get(paste0("lpi",dfi$vlist))
  p <- lp[[dfi$vnames]]
  vpath <- tempfile(fileext = ".png")
  f_ggsave(vpath = vpath,
           pplot=p,
           vw = 6,
           vh = 5)
  return(vpath)
})

limg <- lapply(lpath,\(li){
    image_read(li) %>%
        image_trim() %>% 
        image_scale(.,"x1416") # mesma altura
})
f_append <- \(vi=0,vstack=FALSE){
  vind <- c(1,2) + 2*vi
  image_append(c(limg[[vind[[1]]]], limg[[vind[[2]]]]), stack = vstack)
}
limg_final <- lapply(0:2,f_append)
f_append <- \(x,y){
  image_append(c(x, y), stack = TRUE)
}
img_final <- Reduce("f_append",limg_final)
image_write(img_final,"figuras_e_relatorios/figura_ABC_completa.png")
# 
library(magick)
f_img <- \(imgpath){
    vimg <- image_read(imgpath) %>%
        image_trim()
    image_write(vimg,imgpath)
    gc()
}
f_img("figuras_e_relatorios/figura_ABC_completa.png")
```


## Figure 3 from main text: comparison between model performance and SaTScan-based risk classification of CIRs


```{r corrplot, eval=TRUE, include=FALSE}
library(corrplot)
library(magick)
# exemplo:
# corrplot(M, method = 'ellipse', order = 'AOE', type = 'upper')
dfp <- df_ad %>% 
    group_by(data) %>% 
    summarise(SIG = sum(casos==1)/63,nSIG=sum(casos==0)/63) %>% 
    pivot_longer(-data) %>% 
    mutate(name = factor(name,
                         levels=c("SIG","nSIG"),
                         labels=c(
                             "RR > 1",
                             "RR ≤ 1")
                         )
           ) %>% 
    filter(name=="RR > 1") %>% 
    rename(propRR = value) %>% 
    select(-name)

dfp_sepi <- inner_join(
    dfwis_sepi %>% select(-n),
    dfsumm2 %>% 
      filter(!grepl("Naive",method)) %>% 
      select(-Sensitivity,-BalancedAccuracy),
    by=c("base"="data")
    ) %>% 
    relocate(method,.after="base") %>% 
    mutate(method=droplevels(method)) %>% 
    rename(TPR=TruePositiveRate,FPR=FalsePositiveRate) %>% 
    select(base:MAE,TPR:FPR,PPV:NPV,Youden) %>% 
    inner_join(y=.,x=dfp,by=c("data"="base"))

f_gsub <- \(vc){
  gsub("^TP$","True Positive (TP)",vc) %>% 
    gsub("^TN$","True Negative (TN)",.) %>% 
    gsub("^FP$","False Positive (FP)",.) %>% 
    gsub("^FN$","False Negative (FN)",.) %>% 
    gsub("TruePositiveRate|TPR","TPR",.) %>% 
    gsub("cob50","Pred. Int. 50%",.) %>% 
    gsub("cob95","Pred. Int. 95%",.) %>% 
    gsub("dispersao","Spread",.) %>% 
    gsub("subprev","Underprediction",.) %>% 
    gsub("superprev","Overprediction",.) %>%
    gsub("PPV","PPV",.) %>%
    gsub("NPV","NPV",.) %>%
    gsub("propRR","Prop. RR > 1",.) %>%
    gsub("FalsePositiveRate|FPR","FPR",.)
}
f_padname <- \(mi){
  colnames(mi) <- f_gsub(colnames(mi))
  rownames(mi) <- f_gsub(rownames(mi))
  return(mi)
}

f_corrplot <- \(vtitle,vw=10,vh=10){
    #
    pred_vars <- c("WIS", "Spread", "Underprediction", "Overprediction", 
                   "Pred. Int. 50%","Pred. Int. 95%","MAE","Prop. RR > 1")
    class_vars <- c("TPR", "FPR","PPV", "NPV","Youden","Prop. RR > 1")
    #
    dfcor <- filter(dfp_sepi,method==vtitle) %>% 
        as.data.frame()
    #
    mcor <- cor(select(dfcor,-data,-method),method = "spearman")
    p_mat <- cor.mtest(select(dfcor,-data,-method), method = "spearman")$p
    #
    mcor <- f_padname(mcor)
    mcor <- mcor[pred_vars,class_vars]
    mcor["Prop. RR > 1","Prop. RR > 1"] <- NA
    p_mat <- f_padname(p_mat)
    p_mat <- p_mat[pred_vars,class_vars]
    p_mat["Prop. RR > 1","Prop. RR > 1"] <- NA
    text_colors <- ifelse(is.na(p_mat) | p_mat >= 0.05, "white", "black")
    #
    vpath <- tempfile(fileext = ".png")
    png(
        filename = vpath, 
        width = vw,
        height = vh,
        units = "in",
        res = 100
        )
    #
    corrplot(mcor, method = "ellipse", order = "original", 
             type = "full", tl.col = "black", tl.srt = 0,
             cl.pos = 'b',na.label = " ",
             addCoef.col = text_colors, number.cex = 0.8)
    #
    dev.off()
    #
    return(vpath)
}
lpath <- lapply(levels(dfp_sepi$method),f_corrplot)
names(lpath) <- levels(dfp_sepi$method)
limg <- lapply(names(lpath),\(li){
  #
  vpath <- lpath[[li]]  
  img <- image_read(vpath) %>%
        image_trim() 
  #
  info <- image_info(img)
  largura <- info$width
  #
  banner_titulo <- image_blank(
    width = largura, 
    height = 80,
  color = "white") %>% 
  image_annotate(
      text = li,
      gravity = "center",
      size = 28,         
      font = "sans",     
      color = "black",   
      weight = 700)
  #
  image_append(c(banner_titulo, img), stack = TRUE) %>% 
    image_trim()
  }
 )
image_write(limg[[2]],
            path = "figuras_e_relatorios/correlation_plot.png")
# img_final <- image_append(do.call("c",limg),stack = FALSE)
# image_write(img_final,path = "figuras_e_relatorios/correlation_plot.png")
#@ não vejo pq fazer por CIR, pois a classificação do SatSCan é contextual.
#@ Já a agregação pela unidade espacial remove o contexto temporal.
dfp_cir <- inner_join(
    dfwis_cir %>% 
        st_drop_geometry() %>% 
        select(cod_space,WIS:subprev,MAE),
    dfsumm_cir %>% 
        st_drop_geometry() %>% 
        select(cod_space,method,contains("Rate")) %>% 
        filter(!grepl("Naive",method))
    ) %>% 
    select(-cod_space) %>% 
    mutate(method=droplevels(method))
```


## Fig with the extra material from the confusion matrix

### By week

```{r}
library(patchwork)
lp <- list()
lp[[1]] <- dfsumm2 %>%
    select(data:FP) %>%
    pivot_longer(cols=FP:FN) %>%
    mutate(name=factor(name,
                       levels=c("TP","TN","FP","FN"),
                       labels=c(
                           "True Positive (TP)",
                           "True Negative (TN)",
                           "False Positive (FP)",
                           "False Negative (FN)"
                           )
                       )
           ) %>%
    ggplot(aes(x=data,y=value,color=method)) +
    geom_point(size=1.12,alpha=0.5) +
    geom_line(linewidth=1.15,alpha=0.5) +
    scale_color_manual("",values=c(
        "Standard Bayesian prediction"="red",
        "Moving average of Bayesian predictions"="blue",
        "Naive baseline"="black")) +
    facet_wrap(~name,nrow=2) + #,scales="free"
    labs(y="",x="") +
    theme_bw() +
    theme(legend.position="bottom",
          legend.margin = margin(t = -4, b = 0, unit = "pt"),
          legend.box.margin = margin(t = -10, b = 0, unit = "pt"),
          legend.box.spacing = unit(0, "pt"),
          strip.background = element_rect(fill="white",colour = "black"))
#
dfp <- dfsumm2 %>% pivot_longer(Sensitivity:F1) %>% 
    mutate(name=factor(name,
                       levels=names(dfsumm2)[c(7:16)])
           ) %>% 
    ungroup()
lp[[2]] <- dfp %>% 
  filter(!grepl("Rate|Sensitivity|F1|Accuracy",name)) %>% 
  # filter(name %in% c("PPV","NPV")) %>% 
    ggplot(aes(x=data,y=value,color=method)) +
    scale_color_manual("",values=c(
        "Standard Bayesian prediction"="red",
        "Moving average of Bayesian predictions"="blue",
        "Naive baseline"="black")) +
    geom_line(alpha=0.4) +
    geom_point(alpha=0.4) +
    theme_light() +
    labs(y="",x="epidemiological week") +
    facet_wrap(~name,ncol=2) +
    scale_y_continuous(expand=c(0,0.01)) +
    theme(legend.position="none",
          strip.text=element_text(color = "black",face = "bold"),
          strip.background = element_rect(fill="white",colour = "black"))
p <- wrap_plots(lp,ncol=1) +
  plot_annotation(tag_levels = 'A')
p

# setwd("./metricas_forescan-main/")
ggsave(filename = "figuras_e_relatorios/metricas_classificacao.png",
       plot=p,
       width = 12,
       height = 20,
       dpi=70)
f_img <- \(imgpath){
    vimg <- image_read(imgpath) %>%
        image_trim()
    image_write(vimg,imgpath)
}
f_img("figuras_e_relatorios/metricas_classificacao.png")
```


### By CIR

```{r graficos de confusão por CIR,include=FALSE,eval=FALSE}
library(ggspatial)
f_plot <- \(vcol="TP"){
  vnamefill <- ifelse(
    vcol%in%c("FN","FP","TN","TP"), "% of weeks", vcol
  )
  if(vcol=="Youden"){
    vlimitsc <- c(-1,1)
    vmidp <- 0
  }else{
    vlimitsc <- c(0,1)
    vmidp <- 0.5
  }
  vtitle <- gsub("TP","True Positive (TP)",vcol) %>% 
    gsub("TN","True Negative (TN)",.) %>% 
    gsub("FP","False Positive (FP)",.) %>% 
    gsub("FN","False Negative (FN)",.) %>% 
    gsub("TruePositiveRate","True Positive Rate (TPR)",.) %>% 
    gsub("FalsePositiveRate","False Positive Rate (FPR)",.)
  dfsumm_cir %>% 
    mutate(
        across(
            .cols = FN:FP,
            .fns = ~.x/34
            ),
        method=factor(method,
                      levels = c("Standard Bayesian prediction",
                                 "Moving average of Bayesian predictions",
                                 "Naive baseline"),
                      labels = c("Standard Bayes",
                                 "MA Bayes",
                                 "Naive baseline"))
        ) %>% 
    ggplot(aes(fill=.data[[vcol]])) +
    geom_sf(color="gray") +
            scale_fill_gradient2(low = "purple", 
                                 mid = "white", 
                                 high = "yellow", 
                                 midpoint = vmidp,
                                 name = vnamefill,
                                 limits = vlimitsc) +
            facet_wrap(~method,nrow=1) +
            # 1. Ajuste dos eixos para não ter sobras (margens zeradas)
            coord_sf(expand = FALSE) +
            # 2. Adiciona a Barra de Escala (canto inferior esquerdo)
            annotation_scale(location = "bl", width_hint = 0.3,style = "ticks") +
            # 3. Adiciona o Norte (canto superior direito)
            annotation_north_arrow(
                location = "tr", 
                which_north = "true",
                pad_x = unit(0.15, "in"), 
                pad_y = unit(0.15, "in"),
                height = unit(0.8, "cm"),  # Reduz o tamanho da seta
                width = unit(0.8, "cm"),
                style = north_arrow_orienteering(
                    fill = c("black", "white"),
                    line_col = "black",
                    text_size = 7)
                ) +
            # 4. Manutenção do tema limpo para mapas
            theme_bw() +
    ggtitle(label=vtitle) +
            theme(
                legend.position="bottom",
                plot.title=element_text(hjust=0.5,face="bold"),
                panel.grid = element_blank(),
                axis.text = element_blank(),
                axis.ticks = element_blank(),
                strip.background = element_rect(fill="white"),
                strip.text = element_text(face = "bold",size=8,color="black"),
                text = element_text(size=10)
            )
}
lp <- lapply(names(dfsumm_cir)[c(5:8,11:13,15)],f_plot)
names(lp) <- names(dfsumm_cir)[c(5:8,11:13,15)]
fggsave <- \(vname){
  pi <- lp[[vname]]
  ggsave(filename = paste0("figuras_e_relatorios/metricas_confusao_por_CIR/",vname,".png"),
         plot=pi,
         dpi=100,
         width=7.2,
         height=7
         )
}
lpath <- lapply(names(lp),fggsave)
lapply(lpath,f_img)
```




```{r graficos por CIR, include=FALSE,eval=FALSE,fig.width=14,fig.height=10}
dfp <- dfsumm_cir %>% pivot_longer(Sensitivity:F1) %>% 
    mutate(name=factor(name,
                       levels=names(dfsumm2)[c(7:16)])
           ) %>% 
    ungroup()
p <- dfp %>% 
  filter(!grepl("Rate",name)) %>% 
  # filter(name %in% c("PPV","NPV")) %>% 
    ggplot(aes(x=data,y=value,color=method)) +
    scale_color_manual("",values=c(
        "Standard Bayesian prediction"="red",
        "Moving average of Bayesian predictions"="blue",
        "Naive baseline"="black")) +
    geom_line(alpha=0.4) +
    geom_point(alpha=0.4) +
    theme_light() +
    labs(y="",x="epidemiological week") +
    facet_wrap(~name,ncol=2) +
    scale_y_continuous(expand=c(0,0.01)) +
    theme(legend.position="none",
          strip.text=element_text(color = "black",face = "bold"),
          strip.background = element_rect(fill="white",colour = "black"))
ggsave(filename=paste0(getwd(),"/fig2.png"),
       plot=p,
       dpi=200,
       width=12,
       height=10)
p
```


## gráfico extra: TPR e FPR ~ #RR>1_real

```{r}
dfp <- df_ad %>% 
    group_by(data) %>% 
    summarise(SIG = sum(casos==1),nSIG=sum(casos==0)) %>% 
    pivot_longer(-data) %>% 
    mutate(name = factor(name,
                         levels=c("SIG","nSIG"),
                         labels=c(
                             "RR > 1",
                             "RR ≤ 1")
                         )
           ) %>% 
    filter(name=="RR > 1") %>% 
    rename(RRmaior1 = value) %>% 
    select(-name)
dfp <- inner_join(
    dfp,
    select(dfsumm2,contains("Rate"),data:method)
    ) %>% 
    pivot_longer(contains("Rate")) %>% 
    mutate(
        name=case_when(
            grepl("True",name) ~ "TPR",
            grepl("False",name) ~ "FPR"),
        name=factor(name,levels=c("TPR","FPR")),
        RRmaior1=RRmaior1/63
        )
dfp %>% 
    ggplot(aes(x=RRmaior1,y=value)) +
    geom_point() +
    geom_smooth(se=FALSE,method="lm") +
    facet_grid(method~name)
```




## Method performance evaluation metrics

```{r resumo-das-metricas,echo=FALSE,eval=FALSE,include=FALSE}
f_across <- \(vx){
    vavg <- mean(vx)
    vsd <- sd(vx)
    paste0(sprintf("%.2f", vavg), " (", sprintf("%.2f", vsd), ")")

}
dftab <- dfsumm2 %>%  
    group_by(method) %>% 
    summarise(
        across(.cols = c(TruePositiveRate,
                         FalsePositiveRate,
                         Specificity,
                         PPV,
                         NPV,
                         Accuracy,
                         #BalancedAccuracy,
                         Youden,
                         F1
                         ),
               .fns = f_across
               ),
        .groups = "drop"
        ) %>% 
    mutate(
        method=case_when(
            grepl("Stand",method) ~ "Std. Bayesian prediction",
            grepl("Movin",method) ~ "MA Bayesian prediction",
            grepl("Naiv",method) ~ "Naive baseline"),
        method=factor(method,
                      levels=c("Std. Bayesian prediction",
                               "MA Bayesian prediction",
                               "Naive baseline"))
        
        )
ft_metricas <- 
  flextable(dftab) |>
  theme_booktabs() |>
  # Tamanho e fonte
  fontsize(size = 14, part = "all") |>
  height(height = 0.8) %>%
font(fontname = "Arial", part = "all") |>
  # Alinhamentos
  align(align = "center", part = "header") |>
  bold(part = "header") |>
    set_header_labels(
    method = "Method",
    TruePositiveRate = "TPR",
    FalsePositiveRate = "FPR",
    Specificity = "Specificity",
    PPV = "PPV",
    NPV = "NPV",
    Accuracy = "Accuracy",
    Youden = "Youden",
    F1 = "F1"
  ) %>% 
  set_table_properties(layout = "autofit") |>
  # set_caption("Summary of classification performance metrics for nowcasting methods. Values = average (Standard Deviation)") %>% 
  #   set_caption(
  #   "Table 1. Summary of classification performance metrics for nowcasting methods.<br>
  #   <span style='font-weight:normal;'>Values represent the mean (standard deviation) of each metric across all epidemiological weeks and CIRs, calculated separately for the standard Bayesian nowcast and the smoothed Bayesian nowcast (three-week moving average). Metrics include: True Positive Rate (TPR, sensitivity), False Positive Rate (FPR, 1 - specificity), Specificity, Positive Predictive Value (PPV), Negative Predictive Value (NPV), Accuracy, Youden's Index, and F1 score. Higher values indicate better performance for all metrics except FPR, where lower values are desirable.</span>"
  # ) %>% 
  autofit()
ft_metricas
save_as_image(ft_metricas, 
              "figuras_e_relatorios/tab1_sumario.png", 
              zoom = 3,
              webshot = list(vwidth = 1500, vheight = 1000)
)
```

```{r}
f_img <- \(imgpath){
    vimg <- image_read(imgpath) %>%
        image_trim()
    image_write(vimg,imgpath)
}
f_img("figuras_e_relatorios/esquema_trabalho.png")

```

